Healthcare Insights and Sociodemographic Analysis of New York: A Comprehensive Analysis of Patient Records in 2013

### 2023-09-30 STAT405Project.qmd
added plot

```{r}
#| echo: false
#install.packages("readr")
#install.packages("xtable")
library(readr)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)

@@

```{r}
#| echo: false
cleaned_data$'Total Costs'[cleaned_data$'Total Costs' > quantile(cleaned_data$'Total Costs', 0.95)] <- quantile(cleaned_data$'Total Costs', 0.95)

summary(cleaned_data$'Total Costs')

boxplot(cleaned_data$'Total Costs', 
        main="Box Plot of Total Costs (Outliers Truncated)",
        ylab="Total Costs",
        horizontal = TRUE)
```

@@

cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
payments <- table(cleaned_data$`Payment Typology 1`)
pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA, )
legend(1,1,1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)

@@

```{r}
#| echo: false
ageGroups <- table(cleaned_data$`Age Group`)
barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
```

@@

heatmap(heatmap_data,
        Rowv = NA,          
        Colv = NA,       
        col = colorRampPalette(c("white", "blue"))(50),  
        xlab = "Gender",    
        main = "Gender v. Age Group",  
        labRow = row_names, 
        labCol = col_names)

@@

```{r}
#| echo: false
#install.packages("ggplot2")
library(ggplot2)
ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
  geom_bar() +
  labs(
    title = "Stacked Bar Chart by Gender and Race",
    x = "Gender",
    y = "Count"
  ) +
  scale_fill_manual(
    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
    name = "Race"
  ) +
  theme_minimal()

@@

```{r warning = FALSE}
#| echo: false
#install.packages("ggplot2")
ggplot(cleaned_data) + 
  aes(x = cleaned_data$'APR Risk of Mortality', y = cleaned_data$"Total Costs", color = "salmon") + 
  geom_boxplot() +
  labs(title = "Total Costs on Risk of Mortality",
       x = "APR Risk of Mortality",  # X-axis title
       y = "Total Costs")  # Y-axis title

@@

@@

```{r warning = FALSE}
#| echo: false
ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`))+
  geom_point(size = 0.2) +
  facet_grid(Race ~ `Type of Admission`) +
  labs(title = "Length of Stay v. Total Charges", x = "Length of Stay", y = "Total Charges")

>>>>

```{r}
#| echo: false
# install.packages("readr")
# install.packages("xtable")
library(readr)
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv", show_col_types = FALSE)

@@

```{r}
#| echo: false
cleaned_data$"Total Costs"[cleaned_data$"Total Costs" > quantile(cleaned_data$"Total Costs", 0.95)] <- quantile(cleaned_data$"Total Costs", 0.95)

summary(cleaned_data$"Total Costs")

boxplot(cleaned_data$"Total Costs",
    main = "Box Plot of Total Costs (Outliers Truncated)",
    ylab = "Total Costs",
    horizontal = TRUE
)
```

@@

cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
payments <- table(cleaned_data$`Payment Typology 1`)
pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA, )
legend(1, 1, 1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)

@@

```{r}
#| echo: false
ageGroups <- table(cleaned_data$`Age Group`)
barplot(ageGroups,
    main = "Age Group Distribution", xlab = "Age Group",
    ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)
)
```

@@

heatmap(heatmap_data,
    Rowv = NA,
    Colv = NA,
    col = colorRampPalette(c("white", "blue"))(50),
    xlab = "Gender",
    main = "Gender v. Age Group",
    labRow = row_names,
    labCol = col_names
)

@@

```{r}
#| echo: false
# install.packages("ggplot2")
library(ggplot2)
ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
    geom_bar() +
    labs(
        title = "Stacked Bar Chart by Gender and Race",
        x = "Gender",
        y = "Count"
    ) +
    scale_fill_manual(
        values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
        name = "Race"
    ) +
    theme_minimal()

@@

```{r warning = FALSE}
#| echo: false
# install.packages("ggplot2")
ggplot(cleaned_data) +
    aes(x = cleaned_data$"APR Risk of Mortality", y = cleaned_data$"Total Costs", color = "salmon") +
    geom_boxplot() +
    labs(
        title = "Total Costs on Risk of Mortality",
        x = "APR Risk of Mortality", # X-axis title
        y = "Total Costs"
    ) # Y-axis title

@@

```{r warning = FALSE}
#| echo: false
install.packages("ggExtra")

set.seed(100)
library(ggplot2)
library(ggExtra)
library(dplyr)
random_samples <- sample_n(cleaned_data, 100, replace = FALSE)

Costs <- random_samples$`Total Costs`
Charges <- random_samples$`Total Charges`

g <- ggplot(random_samples, aes(Costs, Charges, howdy)) +
    geom_count(size = 1) +
    geom_smooth(method = "lm", se = F, color = "purple") +
    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
    ggtitle("Total Costs vs Total Charges for 100 Patients")

ggMarginal(g, type = "histogram")

@@

```{r warning = FALSE}
#| echo: false
ggplot(cleaned_data, aes(x = cleaned_data$`Length of Stay`, y = cleaned_data$`Total Charges`)) +
    geom_point(size = 0.2) +
    facet_grid(Race ~ `Type of Admission`) +
    labs(title = "Length of Stay v. Total Charges", x = "Length of Stay", y = "Total Charges")
